---
title: Data 607 Project-Covid Data
author: Alex Moyse
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<H1>The library load in </H1>

```{r}
library(tidyverse)
library(openxlsx)
library(dplyr)
library(zoo)
```


<H1> The Data Grab </H1>

```{r, echo=TRUE}
url.data <- "https://data.cdc.gov/api/views/9mfq-cb36/rows.csv?accessType=DOWNLOAD"
raw <-read.csv(url.data)
raw
```

<H1> Data Engineering </H1>

<H2> Let's Start with Checking the State list </H2>

lets test this?

```{r}
sort(unique(raw$state))

```

<H2> Review </H2>

So firstly FSM is a small island Nation, NYC is New York City, which should be combined with New York State. In addition not all states are included in this list, indicating that not all states provided data. Let's see what happens when we look into the data some more. Just going off the back of my hand and a quick skim of the data list, "AS" "NYC" "PR"  "PW"  "RI"  "RMI" "FSM" "MP" "VI" all appear to be not valid options. First things first, I want to create a new dataframe linking Created_at and submission date, mostly because I am interested in seeing it. 

```{r}
deltaDate <- subset(raw, select = c(submission_date,created_at) )
deltaDate$date_diff <- as.Date(as.character(deltaDate$submission_date), format="%m/%d/%Y")-
                  as.Date(as.character(deltaDate$created_at), format="%m/%d/%Y %H:%M:%S")
deltaDate <- deltaDate[order(deltaDate$date_diff),]
deltaDate
write.csv(deltaDate,"Covid_data_reporting_delta.csv", row.names = FALSE)
```

<H1> Let's Address unique states </H1>

Firstly, there are more entries than there are actual states. At just a first glance over, NYC should be incorporated into NY as its a city in a state. 
 "AK"  "AL"  "AR"   "AZ"  "CA"  "CO"  "CT"  "DC"  "DE"  "FL" "GA"   "HI"  "IA"  "ID"  "IL"  "IN" "KY"  "LA"  "MA"  "MD"  "ME" "MI"  "MN"  "MO"  "MS"  "MT"  "NC"  "ND"  "NE"  "NH"  "NJ"  "NM"  "NV"  "NY" "OH"  "OK"  "OR"  "PA"  "RI"   "SC"  "SD" "TN"  "TX"  "UT"  "VA"   "VT"  "WA"  "WI"  "WV"  "WY" 


```{r}
acceptable_states <- c('AK','AL','AR','AZ','CA','CO','CT','DC','DE','FL','GA','HI','IA','ID','IL','IN','KY','KS','LA','MA','MD','ME','MI','MN','MO','MS','MT','NC','ND','NE','NH','NJ','NM','NV','NY','OH','OK','OR','PA','RI','SC','SD','TN','TX','UT','VA','VT','WA','WI','WV','WY')
all_states <- unique(raw$state)
all_states[!(all_states %in% acceptable_states)]
```

```{r}
data_new <- raw %>%                               # Replacing values
  mutate(state = replace(state, state == 'NYC', 'NY'))
data_new
data_new <- subset(data_new, select = -c(created_at, consent_cases, consent_deaths) )
unique(data_new$state)
data_new


```
```{r}
data_new %>%
  group_by(submission_date, state) %>%
  summarise_all(sum)
```


<H1> Let's Define the quarters </H1>

I really didn't like the approach of aggregating by season as it fails to take into account changes in the COVID situation due to time of year, 


Spring-March 1st to May 31
Summer-June 1st to August 31st
Fall-Sept 1 to November 30th
Winter-Dec 1 to Feb 28/29

```{r}
data_new$yearquarter = as.yearqtr(data_new$submission_date, format = "%m/%d/%Y")
data_new2 <- subset(data_new, select = -c(submission_date, state) )
data_new3 <- data_new2 %>%
  group_by(yearquarter) %>%
  summarise_all(sum)
data_new3
write.csv(data_new3,"Covid_data_year_quarter.csv", row.names = FALSE)
```

<H1> Conclusion </H1>

So we have our data saved and easily accessible. It is ready to be read into anything else, or picked up in R to analyze!

